### 2016-09-11 Project1/article/article.tex
Wrote out most of the theory

\documentclass[10pt, a4paper]{amsart}

\usepackage[]{graphicx}
\usepackage[]{hyperref}
\usepackage[]{physics}

\usepackage[T1]{fontenc}

\title[Solving the Poisson-equation in one dimension]{Solving the Poisson-equation in one dimension \\
  \hrulefill\small{ FYS3150: Computational Physics }\hrulefill}

\author[Sebastian G. Winther-Larsen]{Sebastian G. Winther-Larsen \\
  \href{https://github.com/gregwinther/FYS3150/}{\texttt{github.com/gregwinther}}}

\begin{document}

\begin{titlepage}
\begin{abstract}
Lorem ipsum dolor sit amet, consectur adipiscing lit. Nullam ut lacus eget lorem...
\end{abstract}
\maketitle
\tableofcontents
\end{titlepage}

\section{Introduction}

\section{Theory}
\subsection{The Poisson Equation}

The Poisson equation is a classical equation from electromagnetism. The electrostatic potential $\Phi$ is generated by a localized charge distribution $\rho(\vb{r})$. In three dimensions the equation reads
\begin{equation}
\laplacian\Phi=-4\pi \rho(\vb{r})
\end{equation}
where $\laplacian$ is the Laplace operator. In three dimensions the Laplace operator can be expressed using spherical coordinates, but in this study I am assuming that $\Phi$ and $\rho$ are spherically symmetric, thus reducing the equation to a one-dimensional problem. only dependent on radius $r$.
\begin{equation}
\laplacian=\frac{1}{r^2}\frac{d}{dr}\left(r^2\frac{d\Phi}{dr} \right)
\end{equation}
By substituting $\Phi(r)=\phi(r)/r$ the Poisson equation is reduced to 
\begin{equation}
\frac{d^2\phi}{dr^2}=-4\pi r\rho(r)
\end{equation}
and by letting $\phi \rightarrow u$ and $r \rightarrow x$ one is left with the very simple equation
\begin{equation}
-u''(x)=f(x)
\end{equation}

The inhomogenous term $f$, or source term, is given by the charge distribution $\rho$ multiplied by $r$ and the constant $-4\pi$. In this study, however, the source term will be $f(x)=100e^{-10x}$ and the results can be compared to the analytical solution $u(x)=1-(1-e^{-10})x-e^{-10x}$.

\subsection{Approximation of the Second Derivative}
In this study the one-dimensional Poisson equation will be solved with Dirichlet boundary conditions by rewrting

\end{document}