### 2016-09-14 Project1/article/article.tex
Scrutiny of program. Algorithm part in article almost done.

\documentclass[10pt, a4paper]{amsart}

\usepackage[]{graphicx}
\usepackage[]{hyperref}
\usepackage[]{physics}
\usepackage[]{listings}
\usepackage[T1]{fontenc}
\usepackage{color}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=c++,                 % the language of the code
  otherkeywords={*,...},           % if you want to add more keywords to the set
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
}

\title[Solving the Poisson-equation in one dimension]{Solving the Poisson-equation in one dimension \\
  \hrulefill\small{ FYS3150: Computational Physics }\hrulefill}

\author[Svalheim \& Winther-Larsen]{Trygve Leithe Svalheim \\
   Sebastian G. Winther-Larsen \\
  \href{https://github.com/gregwinther/FYS3150/}{\texttt{github.com/gregwinther}}}

\begin{document}

\begin{titlepage}
\begin{abstract}
Lorem ipsum dolor sit amet, consectur adipiscing lit. Nullam ut lacus eget lorem...
\end{abstract}
\maketitle
\tableofcontents
\end{titlepage}

\section{Introduction}

\section{Theory}
\subsection{The Poisson Equation}

The Poisson equation is a classical equation from electromagnetism. The electrostatic potential $\Phi$ is generated by a localized charge distribution $\rho(\vb{r})$. In three dimensions the equation reads
\begin{equation}
\laplacian\Phi=-4\pi \rho(\vb{r})
\end{equation}
where $\laplacian$ is the Laplace operator. In three dimensions the Laplace operator can be expressed using spherical coordinates, but in this study I am assuming that $\Phi$ and $\rho$ are spherically symmetric, thus reducing the equation to a one-dimensional problem. only dependent on radius $r$.
\begin{equation}
\laplacian=\frac{1}{r^2}\frac{d}{dr}\left(r^2\frac{d\Phi}{dr} \right)
\end{equation}
By substituting $\Phi(r)=\phi(r)/r$ the Poisson equation is reduced to 
\begin{equation}
\frac{d^2\phi}{dr^2}=-4\pi r\rho(r)
\end{equation}
and by letting $\phi \rightarrow u$ and $r \rightarrow x$ one is left with the very simple equation
\begin{equation}
-u''(x)=f(x) \label{eq:2nd}
\end{equation}

The inhomogenous term $f$, or source term, is given by the charge distribution $\rho$ multiplied by $r$ and the constant $-4\pi$. In this study, however, the source term will be $f(x)=100e^{-10x}$ and the results can be compared to the analytical solution $u(x)=1-(1-e^{-10})x-e^{-10x}$.

\subsection{Approximation of the Second Derivative}
In this study the one-dimensional Poisson equation will be solved with Dirichlet boundary conditions by rewriting it as a set of linear equations. The discretized approximation of $u$ is defined as $v_i$ with grid points $x_i=ih$, step size of $h=\frac{1}{n+1}$, in the interval $x_0=0$ to $x_{n+1}=1$ and with boundary conditions $v_0=v_n+1=0$. The interior solution $v_i \forall i \in {1,...,n}$ is to be found. The second order derivative is approximated with the three point formula such that equation \ref{eq:2nd} becomes
\begin{equation}
-\frac{v_{i+1}-2v_i+v{i-1}}{h²}=f_i \label{eq:2approx}
\end{equation}

By defining $\vb{f}=h^2f_i$ one can rewrite equation
\ref{eq:2approx} as $-v_{i+1}-2v_i+v_{i-1}=h^2f_i$. If we ignore the
end points, $i=0$ and $i=n+1$, this equation can be represented as a
matrix equation.
\begin{equation}
\label{eq:matrix}
A\vb{v}=\vb{f}
\end{equation}
\begin{equation}
\begin{bmatrix}
2 & -1 & 0 & 0 & \cdots & 0 & 0 & 0 \\
-1 & 2 & -1 & 0 & \cdots & 0 & 0 & 0 \\
0 & -1 & 2 & -1 & \cdots & 0 & 0 & 0 \\ 
& & \vdots &  & \ddots &  & \vdots & \\
0 & 0 & 0 & 0 & \cdots & -1 & 2 & -1 \\
0 & 0 & 0 & 0 & \cdots & 0 & -1 & 2 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
\vdots \\
v_{n-1} \\
v_n
\end{bmatrix}=
\begin{bmatrix}
f_1 \\
f_2 \\
f_3 \\
\vdots \\
f_{n-1} \\
f_n
\end{bmatrix}
\end{equation}

\section{Algorithms}
Two main methods are implemented and compared. The first method is
gaussian elimination of the tridiagonal matrix $A$, also known as the
\emph{Thomas Algorithm} \cite{thomasalgo}. This is a simplified form of Gaussian
elimination that can ve used to solve tridiagonal systems of
equations. The method is improved upon in order to take into account
the fact that the matrix we are dealing with has the same numbers
along the diagonals. The second method is the LU-decomposistion
method.

\subsection{Tridiagonal Matrix Algorithm}
Our tridiagonal system can be represented by 
\begin{equation}
a_ix_{i-1} + b_ix_i + c_ix_{i+1} = \vb{f}
\end{equation}
with $a_i=-1,b_i=2$ and $c_i=-1$, except for $a_1=0$ and $c_n=0$. Row
reducing a matrix will reveal how the algorithm functions. Limiting
the problem to four dimensions for easier reading and to save the
rainforest\footnote{Writing out a general case will also take up more paper
space}.

\begin{equation}
\left[
\begin{array}{cccc|c}
b_1 & c_1 & 0 & 0 & f_1 \\
a_2 & b_2 & c_2 & 0  & f_2 \\
0 & a_3 & b_3 & c_3 & f_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}
\right] \sim
\left[
\begin{array}{cccc|c}
1 & c_1/b_1 & 0 & 0 & b_1/b_1 \\
0 & b_2-\frac{c_1}{b_1}a_2 & c_2 & 0  & f_2-\frac{f_1}{b_1}a_2 \\
0 & a_3 & b_3 & c_3 & f_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}
\right]
\end{equation}
Now let $\beta_1=b_1$ and $\beta_2=b_2-\frac{c_1}{b_1}a_2$. As new
elements start to appear in vector $\vb{f}$, right to the vertical bar
in the augmented matrix, they are also relabeled to $\tilde{f}_i$. For
example $\tilde{f}_1=f_1/\beta_1$. One more iteration will reveal the
pattern of the algorithm.

\begin{align}
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & (f_2-\tilde{f}a_2)/\beta_2 \\
0 & a_3 & b_3 & c_3 & f_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}
\right] &=
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & a_3 & b_3 & c_3 & f_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}
\right] \\ \sim 
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & 0 & b_3-\frac{c_2}{\beta_2}a_3 & c_3 & f_3-\tilde{f}_2a_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}
\right] &\sim 
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & 0 & 1 & c_3/\beta_3 & (f_3-\tilde{f}_2a_3)/\beta_3 \\
0 & 0 & a_4 & b_4 & f_4
\end{array}\right] \\ 
\sim &\dots \sim
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & 0 & 1 & c_3/\beta_3 & \tilde{f}_3 \\
0 & 0 & 0 & 1 & \tilde{f}_4
\end{array}
\right]
\end{align}

This pattern can be summarized nicely by the following
difference equations. This computation will from now on be referred to
as the ``forward substitution''. 
\begin{equation}
\label{eq:forward1}
\beta_i=b_i-\frac{a_ic_{i-1}}{\beta_{i-1}}, \quad
\tilde{f}_i=f_i-a_i\tilde{f}_{i-1}/\beta_{i-1}, \quad i \in [2,n], \quad
\end{equation}

In c++, this part of the algorithm is implemented just below. Notice
that $b_i$ is overwritten instead of initializing a seperate vector
for $\beta$. 
\lstinputlisting[language=c++, firstline=65, lastline=71]{../problems.cpp}

The Gaussian elimination is not fully complete until a ``backward
substitution'' has been performed as well. This ensures that all
eliments in the row reduced matrix are pivot elements.

\begin{equation}
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & 0 & 1 & c_3/\beta_3 & \tilde{f}_3 \\
0 & 0 & 0 & 1 & \tilde{f}_4
\end{array}
\right] \sim
\left[
\begin{array}{cccc|c}
1 & c_1/\beta_1 & 0 & 0 & \tilde{f}_1 \\
0 & 1 & c_2/\beta_2 & 0  & \tilde{f}_2 \\
0 & 0 & 1 & 0 & \tilde{f}_3 -\frac{c_3}{\beta_3}\tilde{f}_4 \\
0 & 0 & 0 & 1 & \tilde{f}_4
\end{array}
\right]
\end{equation}

The notation is updated again to $v_3 =\tilde{f}_3 -
\frac{c_3}{\beta_3}\tilde{f}_4$. By continuing this iterative process
all the was to the beginning of the matrix by backward substitution
leaves us with a second recurrence relation.
\begin{equation}
\label{eq:backward1}
v_i=(\tilde{f}_i - c_iv_{i+1})/\beta_i, \quad i \in \{n-1,n-2,\dots,1\}
\end{equation}

The backward substitution is implemented in C++ like so:
\lstinputlisting[language=c++, firstline=73,
lastline=77]{../problems.cpp}

An ever-important aspect of any algorithm is how efficient the algorithm is. It
is therefor worthwhile to consider how many floating point operations
the algorithm requires\footnote{The number of $+$,$-$,$*$ and $/$
  operations are counted}. For the forward substitution there are $6$
FLOPS in each iteration and $n-1$ iterations, resulting in $6(n-1)$
FLOPS for the entire loop. Additionally, $1(n-1)$ FLOPS are required
for the backward substitution, adding up to 
\begin{equation}
N_{tridiag}=8(n-1)=\mathcal{O} (8n).
\end{equation}

\subsubsection{Optimizing the algorithm}

Our case is special, because the elements in each of the diagonal arrays in the
tridiagonal matrix, $A$, are equal. $a_i=-1,b_i=2$ and $c_i=-1$,
except for $a_1=0$ and $c_n=0$. This is an advantage, because the
algorithm can be adjusted, and the number of FLOPS consequently reduced.

The algorithm for the forward substitution, in equation
\ref{eq:forward1}, can be simplified to
\begin{equation}
\beta_i= b_i - 1/\beta_{i-1} \quad \tilde{f}_i=f_i-\tilde{f}_{i-1}/\beta_{i-1}
\end{equation}
and the backward substitution, in equation \ref{eq:backward1}, can be
simplified to
\begin{equation}
v_i = (\tilde{f}_i+v_{i+1})/\beta_i
\end{equation}

This results in some new and sleeker C++ code
\lstinputlisting[language=c++, firstline=146,
lastline=158]{../problems.cpp}

One can see that this new and improved algorithm only requires
$6(n-1)$ FLOPS. In general, a Gaussian row reduction algorithm
requires $\frac{2}{3}n^3+\mathcal{O}(n^2)$ FLOPS \cite{morten}, a
great improvement!

\subsection{LU Decomposition}

The LU decompositon method is a method where the matrix $A$ in
equation \ref{eq:matrix} ($A\vb{v}=\vb{f}$) can be rewritten as a product of two matrices
$L$ and $U$, where $L$ is lower triangular and and $U$ is upper
triangular. This method only works for non-singular, invertible
matrices, which the matrix $A$ certainly is.

\section{Results}

\section{Discussion}

\section{Conclusion}

\begin{thebibliography}{10}
  \bibitem{thomasalgo}{Thomas, L.H. (1949), \emph{Elliptic
        Problems in Linear Differential Equations over a
        Network}. Watson Sci. Comput. Lab Report, Columbia University,
      New York.}
    \bibitem{morten}{Hjorth-Jensen, M. (2016). \emph{Computational
        Physics - Lecture Notes 2016}. University of Oslo}
\end{thebibliography}

\end{document}