### 2016-10-20 Project3/article/article.tex
Changed folder structure: article, date etc. Included stuff from trygvels

\documentclass[10pt, a4paper]{amsart}

\usepackage[]{graphicx}
\usepackage[]{hyperref}
\usepackage[]{physics}
\usepackage[]{listings}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}

\lstset{
	frame = single,
	language = C++,
	showstringspaces = false,
	tabsize = 2,
	otherkeywords = {self},
	keywordstyle = \color{blue},
	identifierstyle=\color{deepgreen},
 	stringstyle=\color{orange},
 	backgroundcolor=\color{mygray}
}

\title[Simulation of Solar System]{Simulation of Solar System Using Velocity Verlet Method \\
  \hrulefill\small{ FYS3150: Computational Physics }\hrulefill}

\author[Winther-Larsen \& Svalheim]{Sebastian G. Winther-Larsen \\ 
Trygve L. Svalheim \\
\href{https://github.com/gregwinther/FYS3150/}{\texttt{github.com/gregwinther}}}

\begin{document}

\begin{titlepage}
\begin{abstract}
Two, three, multi. With Euler-Cromer and then Velocity Verlet
\end{abstract}
\maketitle
\tableofcontents
\end{titlepage}

\section{Introduction}

\section{Theory}

\section{Algorithm}

\section{Results}

\section{Discussion}

\section{Conclusion}

\end{document}